---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)
```

```{=html}
<style>
body {
  background-color: #7CB9E8;
  color: #000000;
  font-family: "Segoe UI", Arial, sans-serif;
}

h2 {
  color: #000000;
  border-bottom: 2px solid #1f4e79;
  padding-bottom: 5px;
}

}
</style>
```

## Cat Daily Behaviour Observation Report

This report tracks my cat’s daily behaviours over time using observational logging data. The purpose of this analysis is to understand patterns in behaviour type, duration, and time of day. Here is the link of the form attached <https://docs.google.com/forms/d/e/1FAIpQLScvb_lot9qNF2asqD_J2QOGHVkKNkuYZU0eIWPEowCanIsjWg/viewform?usp=dialog>

### data importing and cleaning

```{r}
library(tidyverse)
logged_data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSLTBoADhg_ojn2rAHurd1spkdrEpFLZl7xr79pZnWksfGewzPKy0uKqeNTal0B5bjhO_tZfS4lvfwl/pub?output=csv")
latest_data <- logged_data %>% rename(timestamp = 1, 
                                      behavior = 2, 
                                      duration = 3, 
                                      location = 4, 
                                      time = 5)

#latest_data
latest_data <- latest_data %>%
  mutate(
    behavior = case_when(
      str_detect(behavior, "Sleeping") ~ "Sleeping",
      str_detect(behavior, "Eating") ~ "Eating",
      str_detect(behavior, "Grooming") ~ "Grooming",
      str_detect(behavior, "meowing") ~ "Meowing",
      TRUE ~ behavior
    )
  )
# data cleaning. Because variable behavior contains emojis (which can affect statistics).
latest_data
# check if it works
```

### Summary of the data

```{r}
mean_duration <- mean(latest_data$duration)
max_duration <- max(latest_data$duration)

most_common_behavior <- latest_data %>%
  count(behavior, sort = TRUE) %>%
  slice(1)

most_common_time <- latest_data %>%
  count(time, sort = TRUE) %>%
  slice(1)

# print out
cat(
  paste0(
    "The average duration of an observed behaviour was ",
    round(mean_duration, 1),
    " minutes.\n"
  ),
  paste0(
    "The longest recorded behaviour lasted ",
    max_duration,
    " minutes.\n"
  ),
  paste0(
    "The most frequently observed behaviour was ",
    most_common_behavior$behavior,
    ", with ",
    most_common_behavior$n,
    " observations.\n"
  ),
  paste0(
    "The most common time of day for observations was ",
    most_common_time$time,
    ", with ",
    most_common_time$n,
    " observations."
  )
)
```

## Bar Chart

### Behaviour frequency

```{r}
ggplot(latest_data, aes(x = behavior)) +
  geom_bar() +
  labs(
    title = "Frequency of Cat Behaviours",
    x = "Behaviour",
    y = "Count"
  ) +
  theme_minimal()
```

### Time period distribution

```{r}
ggplot(latest_data, aes(x = time)) +
  geom_bar() +
  labs(
    title = "Observations by Time of Day",
    x = "Time of Day",
    y = "Count"
  ) +
  theme_minimal()
```

### Average duration of behavior

```{r}
avg_data <- latest_data %>%
  group_by(behavior) %>%
  summarise(avg_time = mean(duration))
ggplot(avg_data, aes(x = behavior, y = avg_time)) +
  geom_col() +
  labs(
    title = "Average Duration of Each Behaviour",
    x = "Behaviour",
    y = "Average Duration (minutes)"
  ) +
  theme_minimal()
```
